\bf Drift Removal for Time Series Data Using Quantile Trend Filtering

### 2019-01-03 Manuscript/detrendify_JCGS.tex
Description of ADMM algorthim for windows, and smoothing splines, function for lambda selection using SIC and validation dataset.
updated and organized manuscript

\documentclass[12pt]{article}
\RequirePackage{amsthm,amsmath,amsbsy,amsfonts}
\usepackage{bm,graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
%\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
%\RequirePackage{hypernat}
\usepackage{paralist, booktabs, multirow}
%\usepackage{url}

\newtheorem{lemma}{Lemma}
\newcommand{\blind}{0}
\newcommand{\trace}{\text{Trace}}
\numberwithin{equation}{section}
\theoremstyle{plain}
%\newtheorem{thm}{Theorem}[section]
%\newtheorem{mydef}{Definition}
%\newtheorem{remark}{Remark}

%% ----------------------------------------------------------------------
%% Packages
%% ----------------------------------------------------------------------
%\usepackage{amsmath,amsfonts,amssymb}

\usepackage{amssymb}
\usepackage[usenames]{color}
%\usepackage{amsbsy,amsthm,amsmath,amssymb}

%% For \mathscr
\usepackage[mathscr]{eucal}

%% For \llbracket and \rrbracket, \varoast, \varoslash
\usepackage{stmaryrd}

%% For \boldsymbol
%\usepackage{amsbsy}

%%% For \bm (bold math)
%\usepackage{bm}

%% For proper spacing in text macros
\usepackage{xspace}

%% For special lists like inparaenum, compactenum, compactitem
\usepackage{paralist}

%% Turn on/off options for PDF
\usepackage{ifpdf}

%% Graphics
\usepackage{graphicx}
\ifpdf
\usepackage{epstopdf}
% REMOVE NEXT LINE TO REVERT TO PNG FILES
\DeclareGraphicsExtensions{.eps,.pdf}
\else
\usepackage{epsfig}
\fi

%% For \subfloat command
\usepackage{subfig}

%% For hyperlinks. Should always be the last package. Omit for DVI version.
\ifpdf
\usepackage[colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\else
\newcommand{\href}[2]{{#2}}
\usepackage{url}
\fi

%% For tabulary
\usepackage{tabulary}

%% Show labels
%\usepackage[notref,notcite]{showkeys}

%% Algorithm
\usepackage{algorithm}
\usepackage{algpseudocode}

%% ----------------------------------------------------------------------
%% Hyper-linked References
%% ----------------------------------------------------------------------

\ifpdf
\newcommand{\Sec}[1]{\hyperref[sec:#1]{Section~\ref*{sec:#1}}} %section
\newcommand{\App}[1]{\hyperref[sec:#1]{Appendix~\ref*{sec:#1}}} %appendix
\newcommand{\Eqn}[1]{\hyperref[eq:#1]{{\rm (\ref*{eq:#1})}}} %equation
\newcommand{\Part}[1]{\hyperref[part:#1]{(\ref*{part:#1})}} %part of theorem
\newcommand{\Fig}[1]{\hyperref[fig:#1]{Figure~\ref*{fig:#1}}} %figure
\newcommand{\Tab}[1]{\hyperref[tab:#1]{Table~\ref*{tab:#1}}} %table
\newcommand{\Thm}[1]{\hyperref[thm:#1]{Theorem~\ref*{thm:#1}}} %theorem
\newcommand{\Lem}[1]{\hyperref[lem:#1]{Lemma~\ref*{lem:#1}}} %lemma
\newcommand{\Prop}[1]{\hyperref[prop:#1]{Proposition~\ref*{prop:#1}}} %proposition
\newcommand{\Cor}[1]{\hyperref[cor:#1]{Corollary~\ref*{cor:#1}}} %corollary
\newcommand{\Def}[1]{\hyperref[def:#1]{Definition~\ref*{def:#1}}} %definition
\newcommand{\Alg}[1]{\hyperref[alg:#1]{Algorithm~\ref*{alg:#1}}} %algorithm
\newcommand{\Ex}[1]{\hyperref[ex:#1]{Example~\ref*{ex:#1}}} %example
\newcommand{\As}[1]{\hyperref[as:#1]{Assumption~{\rm\ref*{as:#1}}}} %assumption
\newcommand{\Reg}[1]{\hyperref[as:#1]{Condition~\ref*{reg:#1}}} %regularity condition
\newcommand{\AlgLine}[2]{\hyperref[alg:#1]{line~\ref*{line:#2} of Algorithm~\ref*{alg:#1}}}
\newcommand{\AlgLines}[3]{\hyperref[alg:#1]{lines~\ref*{line:#2}--\ref*{line:#3} of Algorithm~\ref*{alg:#1}}}
\else
\newcommand{\Sec}[1]{{Section~\ref{sec:#1}}} %section
\newcommand{\App}[1]{{Appendix~\ref{sec:#1}}} %appendix
\newcommand{\Eqn}[1]{{(\ref{eq:#1})}} %equation
\newcommand{\Part}[1]{{(\ref{part:#1})}} %part of proof
\newcommand{\Fig}[1]{{Figure~\ref{fig:#1}}} %figure
\newcommand{\Tab}[1]{{Table~\ref{tab:#1}}} %table
\newcommand{\Thm}[1]{{Theorem~\ref{thm:#1}}} %theorem
\newcommand{\Lem}[1]{{Lemma~\ref{lem:#1}}} %lemma
\newcommand{\Prop}[1]{{Property~\ref{prop:#1}}} %property
\newcommand{\Cor}[1]{{Corollary~\ref{cor:#1}}} %corollary
\newcommand{\Def}[1]{{Definition~\ref{def:#1}}} %definition
\newcommand{\Alg}[1]{{Algorithm~\ref{alg:#1}}} %algorithm
\newcommand{\Ex}[1]{{Example~\ref{ex:#1}}} %example
\newcommand{\As}[1]{{Assumption~\ref{as:#1}}} %assumption
\newcommand{\Reg}[1]{{R~\ref{reg:#1}}} %regularity condition
\newcommand{\AlgLine}[2]{{line~\ref{line:#2} of Algorithm~\ref{alg:#1}}}
\newcommand{\AlgLines}[3]{{lines~\ref{line:#2}--\ref{line:#3} of Algorithm~\ref{alg:#1}}}
\fi

\usepackage{paralist, booktabs, multirow}

%% ----------------------------------------------------------------------
%% Environments
%% ----------------------------------------------------------------------

%% Assumptions environments
\newtheorem{assumption}{Assumption}[section]
%% Special proof where box is inside final equation
% The function \proof and \endproof are defined inside the SIAM class file.
% Be sure to use \myproofend inside the final equation to produce the
% special box that signifies the end of the proof.
\newenvironment{myproof}{\proof}{}
\newcommand{\myproofend}{\quad\endproof}
\newtheorem{theorem}{Theorem}[section]

%Added by Arvind
\usepackage[normalem]{ulem}
\newcommand{\arvind}[1]{{\color{red} #1}}

%%%%%

%\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
%
%\RequirePackage[OT1]{fontenc}
%\RequirePackage{amsthm,amsmath}
%\RequirePackage [round,authoryear]{natbib}
%\RequirePackage{hypernat}
%
%\usepackage{amsbsy,amsthm,amsmath,amssymb}
%\usepackage{graphicx,booktabs}
%%\usepackage{subfig}
\usepackage{mathtools}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{tikz}
\usetikzlibrary{arrows}
%\setlength{\oddsidemargin}{.75in}
%\setlength{\evensidemargin}{.75in}
%\setlength{\textwidth}{5in}
\newtheorem{proposition}{Proposition}[section]
%\newtheorem{corollary}{Corollary}[section]
%\newtheorem{example}{Example}[section]
%\newtheorem{definition}{Definition}[section]

%% ----------------------------------------------------------------------
%% Definitions
%% ----------------------------------------------------------------------

%% Shorthand for Real
\newcommand{\Real}{\mathbb{R}}
\newcommand{\dom}{{\bf dom}\,}

%% Math notation
\newcommand{\Tra}{^{\sf T}} % Transpose
\newcommand{\Inv}{^{-1}} % Inverse
\newcommand{\tr}{\operatorname{tr}} % Trace
\newcommand{\Ker}{\operatorname{Ker}} % Kernel
\def\vec{\mathop{\rm vec}\nolimits}
\def\mtcz{\mathop{\rm mat}\nolimits} % Matricize operation
\newcommand{\amp}{\mathop{\:\:\,}\nolimits}
\newcommand{\bic}{\text{BIC}}

%% Vectors
\newcommand{\V}[1]{{\bm{\mathbf{\MakeLowercase{#1}}}}} % vector
\newcommand{\VE}[2]{\MakeLowercase{#1}_{#2}} % vector element
\newcommand{\Vbar}[1]{{\bm{\M{A}r \mathbf{\MakeLowercase{#1}}}}} % vector
\newcommand{\Vhat}[1]{{\bm{\hat \mathbf{\MakeLowercase{#1}}}}} % vector
\newcommand{\Vtilde}[1]{{\bm{\tilde \mathbf{\MakeLowercase{#1}}}}} % vector
\newcommand{\Vn}[2]{\V{#1}^{(#2)}} % n-th vector
\newcommand{\VnE}[3]{{#1}^{(#2)}_{#3}} % n-th vector
\newcommand{\VtildeE}[2]{\tilde{\MakeLowercase{#1}}_{#2}} % vector element

%% Matrices
\newcommand{\M}[1]{{\bm{\mathbf{\MakeUppercase{#1}}}}} % matrix
\newcommand{\ME}[2]{\MakeLowercase{#1}_{#2}} % matrix element
\newcommand{\MC}[2]{\V{#1}_{#2}}
\newcommand{\Mr}[2]{\V{#1}_{#2}} % matrix row
\newcommand{\Mhat}[1]{{\bm{\hat \mathbf{\MakeUppercase{#1}}}}} % matrix
\newcommand{\Mtilde}[1]{{\bm{\tilde \mathbf{\MakeUppercase{#1}}}}} % matrix
\newcommand{\MhatC}[2]{\Vhat{#1}_{#2}} % matrix column
\newcommand{\Mbar}[1]{{\bm{\M{A}r \mathbf{\MakeUppercase{#1}}}}} % matrix
\newcommand{\MbarC}[2]{\Vbar{#1}_{#2}} % matrix column
\newcommand{\Mn}[2]{\M{#1}^{(#2)}} % n-th matrix
\newcommand{\Mbarn}[2]{\Mbar{#1}^{(#2)}} % n-th matrix
\newcommand{\Mtilden}[2]{\Mtilde{#1}^{(#2)}} % n-th matrix
\newcommand{\MnTra}[2]{\M{#1}^{(#2){{\sf T}}}} % n-th matrix transpose
\newcommand{\MnE}[3]{\MakeLowercase{#1}^{(#2)}_{#3}} % n-th matrix element
\newcommand{\MnC}[3]{\V{#1}^{(#2)}_{#3}} % n-th matrix column
\newcommand{\MbarnC}[3]{\Vbar{#1}^{(#2)}_{#3}} % n-th matrix column
\newcommand{\MnCTra}[3]{\V{#1}^{(#2){{\sf T}}}_{#3}} % n-th matrix column transpose
\newcommand{\MCTra}[2]{\V{#1}^{{\sf T}}_{#2}} % n-th matrix column transpose

%% Tensors
\newcommand{\T}[1]{\boldsymbol{\mathscr{\MakeUppercase{#1}}}} %tensor
\newcommand{\Tbar}[1]{\boldsymbol{\M{A}r \mathscr{\MakeUppercase{#1}}}} %tensor
\newcommand{\That}[1]{\boldsymbol{\hat \mathscr{\MakeUppercase{#1}}}} %tensor
\newcommand{\Ttilde}[1]{\boldsymbol{\tilde \mathscr{\MakeUppercase{#1}}}} %tensor
\newcommand{\TE}[2]{\MakeLowercase{#1}_{\MI{#2}}} % tensor element with multi-index

%% Matrix-matrix operations
\newcommand{\Kron}{\otimes} %Kronecker
\newcommand{\Khat}{\odot} %Khatri-Rao
\newcommand{\Hada}{\ast} %Hadamard
\newcommand{\Divide}{\varoslash}

% Matricize
\newcommand{\Mz}[2]{\M{#1}_{(#2)}} % n-th mode matricize
\newcommand{\Mzsets}[3]{\TM{#1}{#2\times#3}} % matricze with just the sets specified
\newcommand{\Mzall}[4]{\TM{#1}{#2\times#3\,:\,#4}} % full matricize specification

%% K-Tensor
\newcommand{\KT}[1]{\left\llbracket #1 \right\rrbracket} % kruskal operator
\newcommand{\KTsmall}[1]{\llbracket #1 \rrbracket} % small kruskal operator
\newcommand{\KG}[1]{\langle #1 \rangle} % kruskal "group"

%% text with quads around it
\newcommand{\qtext}[1]{\quad\text{#1}\quad}

%% Derivative
\newcommand{\FD}[2]{\frac{\partial #1}{\partial #2}}

%% Prox-operator
\newcommand{\prox}[2]{\operatorname{prox}_{#1}({#2})}

%% Sign-operator
\newcommand{\sign}[1]{\operatorname{sign}({#1})}

%% convex hull
\newcommand{\conv}[1]{\operatorname{conv}({#1})}

%% For comments to each other
\newcommand{\Remark}[2]{{\color{red}Remark from #1: #2}\xspace}
\newcommand{\Note}[1]{{{\color{red}{#1}}}\xspace}

%\startlocaldefs
%\numberwithin{equation}{section}
%\theoremstyle{plain}
%\newtheorem{thm}{Theorem}[section]
%\endlocaldefs

\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

\pdfminorversion=4

%% ----------------------------------------------------------------------
%% Main Document
%% ----------------------------------------------------------------------
\begin{document}

@@

\if0\blind
{
  \title{\bf Drift Removal for Time Series Data Using Quantile Trend Filtering}
  \author{Halley Brantley\thanks{
    Department of Statistics, North Carolina State University, Raleigh, NC 27695-8203 (E-mail: hlbrantl@ncsu.edu)} \,
         Joseph Guinness\thanks{
    Department of Statistics, North Carolina State University, Raleigh, NC 27695-8203 (E-mail: jsguinne@ncsu.edu)} \,
    and
%    and
    Eric C. Chi\thanks{Department of Statistics, North Carolina State University, Raleigh, NC 27695-8203 (E-mail: eric$\_$chi@ncsu.edu).}    \\}
    \date{}
  \maketitle
} \fi

@@

\bigskip
\begin{abstract}
Abstract
\end{abstract}

\noindent%
{\it Keywords:}  Key words
\vfill

\newpage
\spacingset{1.45} % DON'T change the spacing!
%% ----------------------------------------------------------------------
%% Introduction
%% ----------------------------------------------------------------------
\section{Introduction}

\subsection{Background}

\begin{itemize}
\item \cite{Oh2011} unified framework for non-parametric quantile regression by approximating check loss function with quadratic loss near zero.
\item \cite{KoenkerNgPortnoy1994} introduce quantile smoothing splines. We need to discuss how our approach is different.
\item \cite{Kim2009} introduce the concept of $\ell_1$-trend filtering.
\item \cite{Tib2014} describes properties of trend filtering using quadratic loss. Shows that trend filtering estimates adapt to the local level of smoothness much better than smoothing splines, and exhibit a remarkable similarity to locally adaptive regression splines. Prove that (with the right choice of tuning parameter) the trend filtering estimate
converges to the true underlying function at the minimax rate for functions whose kth derivative is of bounded variation.
\item \cite{Ning2014} address problem of estimating a smooth baseline in noisy data with drift.
\item \cite{Takeuchi2006nonparametric} Nonparametric quantile regression using SVM with Gaussian RBF kernels and check (pinball) loss. 
\item \cite{yuan2006gacv} Comparison of cross-validation methods for quantile smoothing splines.  
\end{itemize}
We propose to use the trend filtering penalty with the check loss function to produce a non-parametric quantile regression estimate that can be computes using a linear time algorithm for removing trends in time series. The formulation was proposed by \cite{Kim2009} as a possible extension of $\ell_1$-trend filtering but not studied. Moreover we extend the basic framework to model multiple quantiles and ensure non-crossing.

\subsection{Application}

	 \begin{figure}[!h] 
	 	\caption{Raw Data - three collocated sensors}
		\includegraphics[width = 5in]{Figures/spodRaw.png}
	\end{figure}
	 \begin{figure}[!h] 
	 	\caption{Detrended data - three collocated sensors}
	\includegraphics[width = 5in]{Figures/spodDetrended.png}
	\end{figure}

Examples:
\begin{itemize}
\item Air quality \citep{Apte2017}
\item ECG \citep{Sanyal2012}
\item Chromatogram baseline estimation  \citep{Ning2014, Ilewicz2015}
\item Galaxy spectrum baseline estimation \citep{Ilewicz2015,Bacher2016}
\item Identifying absorption dips from black body radiation
\end{itemize}

Need to decide between detrending versus detrending + denoising. The BEADS \citep{Ning2014} method does both. We may wish to focus on detrending and then use wavelet SURE denoising as a postprocessing step, i.e. do a two-stage procedure, both of which can be done in linear time. On paper this should be faster than the BEADS procedure. Or we may just want to stick with detrending.  There's Matlab code for BEADS, and the BEADS paper also points to two other popular methods in chromotography.

Things to do:
\begin{itemize}
	\item Convergence of the algorithm
	\item Convergence rate \citep{He2012, He2015, Davis2017}
	\item Timing experiments of LP versus Spingarn
	\item Make an R package - detrendr
	\item Compare on synthetic data quality of solution with existing methods
	\item Do comparisons on real data examples
\end{itemize}

%% ----------------------------------------------------------------------
%% Quantile Regression
%% ----------------------------------------------------------------------
\section{Quantile Regression}

The classic least squares regression is notoriously sensitive to outliers. One remedy to blunt the influence of outliers is to compute the least absolute deviations (LAD) solution in place of the least squares one. Given a design matrix $\M{X} \in \Real^{n \times p}$ and continuous responses $\V{y} \in \Real^n$, we estimate a regression vector $\V{\theta} \in \Real^p$ so that $\M{X}\V{\theta}$ is a good approximation of $\V{y}$. The LAD estimator is a solution to the problem
\begin{eqnarray*}
\underset{\V{\theta}}{\min}\; \frac{1}{n}\lVert \V{y} - \M{X}\V{\theta} \rVert_1.
\end{eqnarray*}
The above optimization problem generalizes the notion of the median of a collection of numbers. A median $\mu$ of $n$ reals $\VE{y}{1}, \ldots, \VE{y}{n}$ is the minimizer of the function
\begin{eqnarray*}
f(u) & = & \frac{1}{n}\sum_{i=1}^n \lvert \VE{y}{i} - \theta \rvert.
\end{eqnarray*}
Recall that the median is the 50th percentile or 0.5-quantile, namely half of the $\VE{y}{i}$ are less than or equal to $\mu$ and the other half is greater than or equal to $\mu$. The median can be generalized to arbitrary $\tau$-quantiles for $\tau \in (0,1)$ to give us quantile regression \citep{Koenker1978}.

First define the so-called ``check function"
\begin{eqnarray*}
\rho_\tau(\Delta) & = & \begin{cases}
\tau \Delta & \text{$\Delta \geq 0$} \\
-(1-\tau)\Delta & \text{$\Delta < 0$}. \\
\end{cases}
\end{eqnarray*}
Then the $\tau$th quantile of the $\VE{y}{i}$ is a minimizer of the function
\begin{eqnarray*}
f_\tau(\theta) & = & \frac{1}{n}\sum_{i=1}^n \rho_\tau(\VE{y}{i} - \theta).
\end{eqnarray*}
Returning to the regression context, we can generalize LAD regression to quantile regression, namely computing the minimizer of the function
\begin{eqnarray*}
f_\tau(\V{\theta}) & = & \frac{1}{n}\sum_{i=1}^n \rho_\tau(\VE{y}{i} - \langle \V{x}_i \mid \V{\theta} \rangle),
\end{eqnarray*}
where $\V{x}_i \in \Real^p$ denotes the $i$th row of $\M{X}$.

%% ----------------------------------------------------------------------
%% Trend Filtering
%% ----------------------------------------------------------------------
\section{Trend Filtering}

In the trend filtering problem \citep{Kim2009, Tib2014}, one is interested in finding an adaptive polynomial approximation to noisy data $\V{y} \in \Real^n$ by solving the following convex problem.
\begin{eqnarray*}
\underset{\V{\theta}}{\arg\min}\; \frac{1}{2n} \lVert \V{y} - \V{\theta} \rVert_2^2 + \lambda \lVert \Mn{D}{k+1}\V{\theta} \rVert_1,
\end{eqnarray*}
where $\lambda \geq 0$ is a regularization parameter that trades off the emphasis on the data fidelity term and the matrix $\Mn{D}{k+1} \in \Real^{(n - k -1) \times n}$ is the discrete difference operator of order $k+1$. To understand the purpose of penalizing $\Mn{D}{k+1}$ consider the difference operator when $k = 0$.
\begin{eqnarray*}
\Mn{D}{1} = \begin{pmatrix}
-1 & 1 & 0 & \cdots & 0 & 0 \\
0 & -1 & 1 & \cdots & 0 & 0 \\
\vdots & & & & & \\
0 & 0 & 0 & \cdots & -1 & 1 \\
\end{pmatrix}
\end{eqnarray*}
Thus, $\lVert \Mn{D}{1}\V{\theta} \rVert_1 = \sum_{i=1}^{n-1} \lvert \VE{\theta}{i} - \VE{\theta}{i+1} \rvert$ which is just total variation denoising in one dimension. The penalty incentivizes solutions which are piece-wise constant. For $k \geq 1$, the difference operator $\Mn{D}{k+1} \in \Real^{(n-k-1) \times n}$ is defined recursively as follows
\begin{eqnarray*}
\Mn{D}{k+1} & = & \Mn{D}{1}\Mn{D}{k}.
\end{eqnarray*}
By penalizing the $k+1$ fold composition of the discrete difference operator, we obtain solutions which are piecewise polynomials of order $k$.

%% ----------------------------------------------------------------------
%% Quantile Trend Filtering
%% ----------------------------------------------------------------------
\section{Quantile Trend Filtering}

We combine the ideas of quantile regression and trend filtering, namely consider the signal approximation problem, where the design $\M{X}$ is the identity matrix.

The estimation of the quantile trend filtering model can be posed as the following optimization problem.
\begin{eqnarray}
\label{eq:quantile_trend}
	\underset{\V{\theta}}{\min}\; \frac{1}{n} \sum_{i=1}^n \rho_\tau(\VE{y}{i} - \VE{\theta}{i}) + \lambda \lVert \Mn{D}{k} \V{\theta} \rVert_1,
\end{eqnarray}
where $\lambda$ is a nonnegative tuning parameter. As with the classic quantile regression, the quantile trend filtering problem can be solved by a linear program. We argue that it is better solved by Spingarn's method of partial inverses.

\section{Related Work}

\begin{itemize}
	\item Quantile splines \citep{Oh2011}
	\item BEADS 
\end{itemize}

%% ----------------------------------------------------------------------
%% Spingarn's method of partial inverses
%% ----------------------------------------------------------------------
\section{Spingarn's method of partial inverses}

We first review Spingarn's method \citep{Spingarn1985}, which solves the following equality constrained convex problem:
\begin{equation}
\label{eq:spingarn}
\begin{split}
\text{minimize}\; & \psi(\V{x}) \\
\text{subject to}\; & \V{x} \in V,
\end{split}
\end{equation}
where $V$ is a subspace. The problem \Eqn{spingarn} can be expressed as the unconstrained optimization problem
\begin{equation}
\label{eq:spingarn2}
\text{minimize}\; \psi(\V{x}) + \iota_V(\V{x}),
\end{equation}
where $\iota_V$ is the indicator function of the set $V$. Spingarn's method applies Douglas-Rachford splitting to the problem \Eqn{spingarn2} to give the following updates.

\begin{eqnarray*}
\Vn{x}{k+1} & = & \prox{t\psi}{\Vn{z}{k}} \\
\Vn{y}{k+1} & = & P_V\left(2\Vn{x}{k+1} - \Vn{z}{k} \right) \\
\Vn{z}{k+1} & = & \Vn{z}{k} + \lambda^{(k)}\left(\Vn{y}{k+1} - \Vn{x}{k+1} \right).
\end{eqnarray*}
The parameter $t$ is a step-size and $\lambda^{(k)}$ is Krasnosel'ski{\u i}-Mann iteration (Need citation). We require that $\lambda^{(k)} \in ]0, 2[$ and that $\sum_n \lambda^{(k)} (2-\lambda^{(k)}) = \infty$. The mapping $P_V$ is the orthogonal projection onto the set $V$.  Note that the algorithm iterates $\Vn{x}{k}$ will converge to a solution to problem $\Eqn{spingarn}$ \citep{Combettes2005}. We need to experiment with different over / under-relaxation parameters $\lambda^{(k)}$ and step sizes $t$. It will converge if we take $\lambda^{(k)} = 1$ and $t = 1$, but we may be able to converge faster in practice by taking non-trivial values.

%% ----------------------------------------------------------------------
%% Applying Spingarn's Method to Quantile Trend Filtering
%% ----------------------------------------------------------------------
\section{Applying Spingarn's Method to Quantile Trend Filtering}

>>>>

\documentclass[12pt]{article}
\RequirePackage{amsthm,amsmath,amsbsy,amsfonts}
\usepackage{graphicx}
%\usepackage{enumerate}
\usepackage{natbib}
\usepackage{url} % not crucial - just used below for the URL

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

\begin{document}

	%\bibliographystyle{natbib}

@@

	\if0\blind
	{
		\title{\bf Drift Removal for Time Series Data Using Quantile Trend Filtering}
		\author{Halley Brantley\thanks{
				Department of Statistics, North Carolina State University, Raleigh, NC 27695 (E-mail: hlbrantl@ncsu.edu)} \,
			Joseph Guinness\thanks{
				Department of Biological Statistics and Computational Biology, Cornell University, Ithaca, NY 14853 (E-mail: guinness@cornell.edu )} \,
			and
			%    and
			Eric C. Chi\thanks{Department of Statistics, North Carolina State University, Raleigh, NC 27695 (E-mail: eric$\_$chi@ncsu.edu).}    \\}
		\date{}
		\maketitle
	} \fi

@@

	\bigskip
	\begin{abstract}
		The text of your abstract.  200 or fewer words.
	\end{abstract}

	\noindent%
	{\it Keywords:}  3 to 6 keywords, that do not appear in the title
	\vfill

	\newpage
	\spacingset{1.5} % DON'T change the spacing!
	\section{Introduction}
	\label{sec:intro}

	We propose to use the trend filtering penalty with the check loss function to produce a non-parametric quantile regression estimate for removing trends in time series. The formulation was proposed by \cite{Kim2009} as a possible extension of $\ell_1$-trend filtering but not studied. Moreover we extend the basic framework to ensure non-crossing while modeling multiple quantiles. We also implement a parallel ADMM algorithm for series that are too large to be computed simultaneously and proposed a modified criteria for choosing the smoothing parameter. We demonstrate through simulation studies that our proposed model provides better or comparable estimates of non-parametric quantile trends than existing methods and is a more effective method of drift removal for low-cost air quality sensors.

	\subsection{Quantile Regression}

	The classic least squares regression is notoriously sensitive to outliers. One remedy to blunt the influence of outliers is to compute the least absolute deviations (LAD) solution in place of the least squares one. Given a design matrix $\mathbf{X} \in \mathbb{R}^{n \times p}$ and continuous responses $y \in \mathbb{R}^n$, we estimate a regression vector $\theta \in \mathbb{R}^p$ so that $\mathbf{X}\theta$ is a good approximation of $y$. The LAD estimator is a solution to the problem
	\begin{eqnarray}
		\underset{\theta}{\min}\; \frac{1}{n}\lVert y - \mathbf{X}\theta \rVert_1.
	\end{eqnarray}
	The above optimization problem generalizes the notion of the median of a collection of numbers. A median $\mu$ of $n$ reals $y_1, \ldots, y_n$ is the minimizer of the function
	\begin{eqnarray}
		f(u) & = & \frac{1}{n}\sum_{i=1}^n \lvert y_i - \theta \rvert.
	\end{eqnarray}
	Recall that the median is the 50th percentile or 0.5-quantile, namely half of the $y_i$ are less than or equal to $\mu$ and the other half is greater than or equal to $\mu$. The median can be generalized to arbitrary $\tau$-quantiles for $\tau \in (0,1)$ to give us quantile regression \citep{Koenker1978}.

	First define the so-called "check function"
	\begin{eqnarray}
		\rho_\tau(\Delta) & = & \begin{cases}
			\tau \Delta & \Delta \geq 0 \\
			-(1-\tau)\Delta & \Delta < 0 \\
		\end{cases}
	\end{eqnarray}
	Then the $\tau$th quantile of the $y_i$ is a minimizer of the function
	\begin{eqnarray}
		f_\tau(\theta) & = & \frac{1}{n}\sum_{i=1}^n \rho_\tau(y_i - \theta).
	\end{eqnarray}
	Returning to the regression context, we can generalize LAD regression to quantile regression, namely computing the minimizer of the function
	\begin{eqnarray}
		f_\tau(\theta) & = & \frac{1}{n}\sum_{i=1}^n \rho_\tau(y_i - \langle x_i \mid \theta \rangle),
	\end{eqnarray}
	where $x_i \in \mathbb{R}^p$ denotes the $i$th row of $\mathbf{X}$.

	%% ----------------------------------------------------------------------
	%% Trend Filtering
	%% ----------------------------------------------------------------------
	\subsection{Trend Filtering}

	In the trend filtering problem \citep{Kim2009, Tib2014}, one is interested in finding an adaptive polynomial approximation to noisy data $y \in \mathbb{R}^n$ by solving the following convex problem.
	\begin{eqnarray}
		\underset{\theta}{\arg\min}\; \frac{1}{2n} \lVert y - \theta \rVert_2^2 + \lambda \lVert \mathbf{D}^{(k+1)}\theta \rVert_1,
	\end{eqnarray}
	where $\lambda \geq 0$ is a regularization parameter that trades off the emphasis on the data fidelity term and the matrix $\mathbf{D}^{(k+1)} \in \mathbb{R}^{(n - k -1) \times n}$ is the discrete difference operator of order $k+1$. To understand the purpose of penalizing $\mathbf{D}^{(k+1)}$ consider the difference operator when $k = 0$.
	\begin{eqnarray}
		\mathbf{D}^{(1)} = \begin{pmatrix}
			-1 & 1 & 0 & \cdots & 0 & 0 \\
			0 & -1 & 1 & \cdots & 0 & 0 \\
			\vdots & & & & & \\
			0 & 0 & 0 & \cdots & -1 & 1 \\
		\end{pmatrix}
	\end{eqnarray}
	Thus, $\lVert \mathbf{D}^{(1)}\theta \rVert_1 = \sum_{i=1}^{n-1} \lvert \theta_i - \theta_{i+1} \rvert$ which is just total variation denoising in one dimension. The penalty incentivizes solutions which are piece-wise constant. For $k \geq 1$, the difference operator $\mathbf{D}^{(k+1)} \in \mathbb{R}^{(n-k-1) \times n}$ is defined recursively as follows
	\begin{eqnarray}
		\mathbf{D}^{(k+1)} & = & \mathbf{D}^{(1)}\mathbf{D}^{(k)}.
	\end{eqnarray}
	By penalizing the $k+1$ fold composition of the discrete difference operator, we obtain solutions which are piecewise polynomials of order $k$.

	%% ----------------------------------------------------------------------
	%% Quantile Trend Filtering
	%% ----------------------------------------------------------------------
	\section{Methods}

	\subsection{Quantile Trend Filtering}

	We combine the ideas of quantile regression and trend filtering, namely consider the signal approximation problem, where the design $\mathbf{X}$ is the identity matrix.

	The estimation of the quantile trend filtering model can be posed as the following optimization problem.
	\begin{eqnarray}
	\label{eq:quantile_trend}
	\underset{\theta}{\min}\; \frac{1}{n} \sum_{i=1}^n \rho_\tau(y_i - \theta_i) + \lambda \lVert \mathbf{D}^{(k)} \theta \rVert_1,
	\end{eqnarray}
	where $\lambda$ is a non-negative tuning parameter. As with the classic quantile regression, the quantile trend filtering problem can be solved by a linear program.

	\subsection{Smoothing parameter choice}
	Our method can easily handle missing data by changing the check loss function to output 0 for missing values. This allows us to leave out validation observations that can be used to select the tuning parameter $\lambda$ and to compare method performance on real data. A number of methods have been proposed for selecting the quantile regression smoothing spline tuning parameter \cite{yuan2006gacv}.  \cite{KoenkerNgPortnoy1994} relate $\lambda$ to the number of interpolated points $p_{\lambda} = \sum I(y_i = \widehat{g}_i(x_i))$, which can be thought of as active knots, they propose the Schwarz criterion for the selection of $\lambda$
	\begin{equation}
	SIC(p_{\lambda}) = log[n^{-1}\sum\rho_{\tau}(y_i - \widehat{g}(x_i))] + \frac{1}{2n}p_{\lambda}log n
	\end{equation}

	The traditional Bayesian Information Criterion (BIC) is given by 
	\begin{equation}
	\mbox{BIC}(s) = -2\log(L\{\hat{\theta}(s)\}) + \nu(s)\log n 
	\end{equation}	
	where $\theta(s)$ is the parameter $\theta$ with those components outside $s$ being set to 0, and $\nu(s)$ is the number of components in $s$. If we assume an asymmetric Laplace likelihood $L(y|\theta) = \left(\frac{\tau^n(1-\tau)}{\sigma}\right)^n\exp\left\{-\sum_i\rho_\tau(\frac{y_i - \theta_i}{\sigma})\right\}$ and the number of non-zero elements of $D\theta$ as $df$
	\begin{equation}
	\mbox{BIC}(df) = 2\sum_i\frac{1}{\sigma}\rho_{\tau}(y_i-\theta_i) + df\log n
	\end{equation} 
	We can choose and $\sigma>0$ and have found empirically that $\sigma =  \frac{1-|1-2\tau|}{2}$ produces stable estimates. \cite{chen2008} proposed the extended BIC for large parameter spaces 
	\begin{equation}
	BIC_{\gamma}(s) = -2\log(L\{\hat{\theta}(s)\}) + \nu(s)\log n  + 2\gamma\log{P\choose j,}~~\gamma \in [0,1]
	\end{equation}
	where $P$ is the total number of possible parameters and $j$ is the number of parameters included in given model. We used this criteria with $\gamma = 1$, $P=n-k$ where $k$ is the order of the differencing matrix and $j = \nu(s)$ is the number of non-zero entries in $D^{(k)}\theta$.

	We used a single dataset to illustrate the difference between the scaled, unscaled and extended BIC criteria.

	\begin{figure}[h!]
	\caption{Degrees of freedom (number of non-zero elements of $D\theta$) by $\log(\lambda)$.} 
	\includegraphics[width = 0.25\linewidth]{Figures/BIC_data.png}
	\includegraphics[width = 0.25\linewidth]{Figures/df_by_lambda.png}
	\includegraphics[width = 0.5\linewidth]{Figures/BIC_by_lambda.png}
	\end{figure}

	\subsection{ADMM for Smoothing Windows}
